---
title: "Women Entrepreneurship and Labor Force"
author: "Adriana Picoral"
date: "4/20/2021"
output: html_document
---

```{r setup, include=FALSE}
library(janitor)
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE)
```

# Data and data question

> What impact do women have on a country's economy?

```{r}
# shortcut for new block cmd+option+i or ctrl+alt+i
# read data in
women_business_data <- read_delim("data/women_in_labor_force.csv",
                                  delim = ";") %>%
  clean_names()
```

# Lead

> How do we measure, in numeric terms, the difference between developed and developing countries?

```{r}
# start with the raw data and then
# map inflation rate to y and level of development to x and add
# geom_boxplot (2:35pm)
women_business_data %>%
  ggplot(aes(x = level_of_development,
             y = inflation_rate)) +
  geom_boxplot() +
  theme_linedraw() +
  labs(y = "inflation rate",
       x = "level of development",
       title = "Distribution of inflation rate",
       subtitle = "across different countries")

```

As we can see in the plot above, developed countries have a lower inflation rate than developing countries.

# Development

> What affects inflation rate? What factor affects a country's economy?

```{r}
# scatter plot of inflation rate entrepreneurship index (geom_point)
women_business_data %>%
  ggplot(aes(x = entrepreneurship_index,
             y = inflation_rate)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_linedraw() +
  labs(y = "inflation rate",
       x = "entrepeneurship index",
       title = "Relationship between inflation rate and entrepeneurship")

```


```{r}
# run linear regression
model_1 <- women_business_data %>%
  lm(formula = inflation_rate ~ entrepreneurship_index) 

# summary of model
summary_model1 <- summary(model_1)

# print out the coefficients
summary_model1$coefficients %>%
  kable()
```

The effect of entrepreneurship on inflation is negative, with inflation going down .13 points for each additional point in entrepreneurship.

# Resolution

> What is the effect of women participanting in their country's economy?

```{r}
women_business_data %>%
  ggplot(aes(y = inflation_rate,
             x = women_entrepreneurship_index)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_linedraw() +
  labs(x = "Women Entreperneurship Index",
       y = "inflation rate",
       title = "Relationship between inflation rate and women entrepreneurship")
```

```{r}
# run linear regression
model_2 <- women_business_data %>%
  lm(formula = inflation_rate ~ women_entrepreneurship_index)

# summary of model
summary_model_2 <- summary(model_2)

# coefficients
summary_model_2$coefficients %>%
  kable()
```

Women entrepreneurship has a statistically significant (p < .05) effect on inflation rate, with each positive point in this index lowering inflation rate by .17 points. This index explains 19% of the variation in the data.

## Overlay both plots

```{r}
ggplot(data = women_business_data) +
  geom_point(aes(y = inflation_rate,
             x = entrepreneurship_index),
             color = "brown4") +
  geom_smooth(aes(y = inflation_rate,
             x = entrepreneurship_index),
             method = "lm", color = "brown4") +
  geom_point(aes(y = inflation_rate,
                 x = women_entrepreneurship_index),
             color = "dodgerblue4") +
  geom_smooth(aes(y = inflation_rate,
                  x = women_entrepreneurship_index),
              method = "lm", color = "dodgerblue4") +
  theme_linedraw() 
  
```



```{r}
# scatterplot
women_business_data %>%
  ggplot(aes(y = inflation_rate,
             x = female_labor_force_participation_rate)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_linedraw() +
  labs(x = "female labor force participation rate",
       y = "inflation rate",
       title = "Relationship between inflation rate and female labor force participation")

# run linear regression
model_3 <- women_business_data %>%
  lm(formula = inflation_rate ~ female_labor_force_participation_rate)

# summary of model_3
summary_model_model_3 <- summary(model_3)

# inspect summary
summary_model_model_3$coefficients %>%
  kable()
```

Female labor force participation rate does not have a significant effect on inflation rate (p > .05).

# Multivariate linear linear

> Does being a member of the European Union have an effect on a countries economy? What about members and non membership and women entreperneurship index?

```{r}
women_business_data %>%
  ggplot(aes(y = inflation_rate,
             x = women_entrepreneurship_index,
             color = european_union_membership)) +
  geom_point() +
  geom_smooth(method = "lm")

# regression
model_4 <- women_business_data %>%
  lm(formula = inflation_rate ~ women_entrepreneurship_index + european_union_membership)

# summary
summary(model_4)
```


